Confluence in Lens Synthesis

### 2020-04-21 papers/BijectiveLenses/IWC/main.tex
add main and citations

  Our proof that we can synthesize all string lenses involves proving a
  confluence-like property, indexed by a an additional binary relation $R$. In
  this model, standard confluence is the specific case where $R$ is equality. We
  show how our proof that the transitive closure of a rewrite system is
  $R$-confluent differs from the proof strategy for standard confluence.
\end{abstract}

@@

To indicate a lens $\Lens$ is well typed, and converts between the languages of
$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$ For synthesis, given $\Regex$ and $\RegexAlt$, we must
find a lens $\Lens$ such that $\Lens : \Regex \Leftrightarrow \RegexAlt$. There
are a number of typing rules for lenses, but the rules for type-equivalence and
composition (shown below) do not permit an efficient search algorithm. Note that
in type equivalence we use a finer notion of equivalence than language
equivalence, we instead use star-semiring equivalence, denoted $\SSREquiv$. We were unable to make
our proof strategy work for language equivalence, so instead guarantee
completeness only on the subset of lenses well-typed with star-semiring
equivalence.
\begin{mathpar}
  \inferrule[\RewriteRegexLensRule{}]
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \SSREquiv \Regex'\\
    \RegexAlt \SSREquiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }

  \inferrule[\ComposeLensRule{}]
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \Regex_2\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3\\
  }
  {
    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Leftrightarrow \Regex_3
  }
\end{mathpar}

Trying to synthesize a lens from the types by applying typing rules ``in
reverse'' gets confounded by these two rules. They require no significant
restriction on the output types, and so can be applied at any time. Furthermore,
applying them in reverse requires pulling regular expressions ``out of thin
air.'' For \RewriteRegexLensRule, these regular expressions need only be
equivalent to the input regular expressions, and for \ComposeLensRule, there are
no restrictions at all!

Our solution is to synthesize in an alternative language of disjunctive normal
form lenses (DNF lenses), typed by disjunctive normal form regular expressions
(DNF REs). Both DNF lenses and DNF REs can be thought of as lenses and regular
expressions, put into a pseudonormal form.

\begin{figure}
  \centering
  \begin{tikzpicture}[auto,node distance=1.5cm]
    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};

    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };

    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};

    \node[text width=2.9cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};

    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};

    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};

    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};

    \path[->] (regex1.east) edge (todnfregex.north west);
    \path[->] (regex2.east) edge (todnfregex.south west);

    \path[->] (todnfregex.north east) edge (dnfregex1.west);
    \path[->] (todnfregex.south east) edge (dnfregex2.west);

    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
    \path[->] (dnfregex2.east) edge (synthdnflens.south west);

    \path[->] (synthdnflens) edge (dnflens);

    \path[->] (dnflens) edge (tolens);

    \path[->] (tolens) edge (lens);
    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
    % Now place a relation (ID=rel1)
    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
    %\node (below-gen) [below=.5cm of sketch-gen] {};
    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
    %\node (below-compl) [below=.5cm of sketch-compl] {};
    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
    %% Draw an edge between rel1 and node1; rel1 and node2
    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);

    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
    %
    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure) (success) ,draw] (surrounding) {};
  \end{tikzpicture}
  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
    \RegexAlt{}, and examples, \Examples{}, are given as input.
    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
    \RegexAlt{}, and \Examples{}.
    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
    that is equivalent to \DNFLens{}.}
  \label{fig:schematic-diagram-synthesis}
\end{figure}

Two typing judgements are involved in typing a DNF lens, $\DNFLens : \DNFRegex
\Leftrightarrow \DNFRegexAlt$ and $\DNFLens \OfRewritelessType \DNFRegex
\Leftrightarrow \DNFRegexAlt$. In both instances, the judgement means $\DNFLens$
is a DNF lens that converts between $\LanguageOf{\DNFRegex}$ and
$\LanguageOf{\DNFRegexAlt}$. If $\DNFLens \OfRewritelessType \DNFRegex
\Leftrightarrow \DNFRegexAlt$, then the typing derivation of $\DNFLens$ never
rewrites the DNF REs. The only typing rule for $\DNFLens : \DNFRegex
\Leftrightarrow \DNFRegexAlt$ is:
\begin{mathpar}
  \inferrule[\RewriteDNFRegexLensRule{}]
  {
    \DNFRegex' \StarOf{\Rewrite} \DNFRegex\\
    \DNFRegexAlt' \StarOf{\Rewrite} \DNFRegexAlt \\
    \DNFLens \OfRewritelessType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt}
  }
  {
    \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex'}{\DNFRegexAlt'}
  }
\end{mathpar}
To type a DNF lens, one first applies a number of rewrites to the DNF REs, then
builds up a DNF lens using rewriteless typing rules. Because of this
phase distinction, equivalent
types only need to be generated once, and composition never need be applied.
This structure corresponds to an efficient algorithm that first searches through type
equivalences with rewrites and then builds up a rewriteless DNF
lens as directed by the syntax of the DNF RE. 
Figure~\ref{fig:schematic-diagram-synthesis} shows the full synthesis
algorithm. The input
regular expressions are converted into DNF form using the
$\ToDNFRegex$ function. Next, a
well-typed DNF lens is synthesized and then converted back into a standard lens
language with the $\ToLens$ function.

Proving this search procedure complete reduces to proving DNF lenses complete
with respect to our standard lens language. In particular, we wish the prove the
following theorem:
\begin{theorem}
  If $\Lens : \Regex \Leftrightarrow \RegexAlt$, then there exists a DNF lens,
  $\DNFLens$, such that $\DNFLens \OfType \ToDNFRegexOf{\Regex} \Leftrightarrow
  \ToDNFRegexOf{\RegexAlt}$ and the semantics of $\Lens$ and $\DNFLens$
  are equivalent.
\end{theorem}

We prove this propety by induction. The particular difficulty lies in the lens
equivalence rule. We begin this case below:

@@

By induction assumption, there exists $\DNFLens \OfType \ToDNFRegexOf{\Regex_1}
\Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}$, where the semantics of $\DNFLens$
are equivalent to those of $\Lens$. By inversion
\begin{mathpar}
  \inferrule
  {
    \ToDNFRegexOf{\Regex_1} \StarOf{\Rewrite} \DNFRegex_1\\
    \ToDNFRegexOf{\RegexAlt_1} \StarOf{\Rewrite} \DNFRegexAlt_1\\
    \DNFLens \OfRewritelessType \DNFRegex_1 \Leftrightarrow \DNFRegexAlt_1
  }
  {
    \DNFLens \OfType \ToDNFRegexOf{\Regex_1} \Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}
  }
\end{mathpar}

@@

We first show that there exists $\DNFRegex_2$ and $\DNFRegexAlt_2$ such that
$\ToDNFRegexOf{\Regex_2} \StarOf{\rightarrow} \DNFRegex_2$ and
$\ToDNFRegexOf{\DNFRegex_1} \StarOf{\rightarrow} \DNFRegex_2$ and
$\ToDNFRegexOf{\RegexAlt_2} \StarOf{\rightarrow} \DNFRegexAlt_2$ and
$\ToDNFRegexOf{\DNFRegexAlt_1} \StarOf{\rightarrow} \DNFRegexAlt_2$. To do this,
we first prove that $\StarOf{\rightarrow}$ is confluent, and that if $\Regex
\SSREquiv \RegexAlt$, then $\ToDNFRegexOf{\Regex}$ is related to
$\ToDNFRegexOf{\RegexAlt}$ through the transitive, reflexive closure of
$\rightarrow$. We then apply a corollary of confluence to show the existance of
such a $\DNFRegex_2$ and $\DNFRegexAlt_2$

@@

After this, we have $\DNFLens \OfRewritelessType \DNFRegex_1 \Leftrightarrow
\DNFRegexAlt_1$ and $\DNFRegex_1 \StarOf{\rightarrow} \DNFRegex_2$ and
$\DNFRegexAlt_1 \StarOf{\rightarrow} \DNFRegexAlt_2$. If we can prove a
confluence-like property that would show the existance of some $\DNFLens'
\OfRewritelessType \DNFRegex_3 \Leftrightarrow \DNFRegexAlt_3$, where
$\DNFRegex_2 \StarOf{\rightarrow} \DNFRegex_3$ and $\DNFRegexAlt_2
\StarOf{\rightarrow} \DNFRegexAlt_3$, we would be done. This property is
$R$-confluence for a properly chosen $R$. This case is diagrammed in Figure~\ref{fig:equiv-case}

@@

>>>>

  Our proof that we can synthesize all string lenses involves proving a
  confluence-like property, indexed by a an additional binary relation $R$. We
  call this property $R$-confluence. In
  this model, standard confluence is the specific case where $R$ is equality. We
  show how our proof that the transitive closure of a rewrite system is
  $R$-confluent differs from the proof strategy for standard confluence.
\end{abstract}

@@

To indicate a lens $\Lens$ is well typed, and converts between the languages of
$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$
For synthesis, given $\Regex$ and $\RegexAlt$, we must find a lens $\Lens$ such
that $\Lens : \Regex \Leftrightarrow \RegexAlt$. In the context of lens
synthesis, there are three sorts of lens typing rules: syntax-directed rules,
composition, and type equivalence.

For syntax-directed rules, the syntax for the types closely mirrors the syntax
of the expressions. As an example, consider the following rule for
disjunction:
\begin{mathpar}
  \inferrule[\OrLensRule{}]
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
    \UnambigOrOf{\Regex_1}{\Regex_2}\\
    \UnambigOrOf{\RegexAlt_1}{\RegexAlt_2}
  }
  {
    \OrLensOf{\Lens_1}{\Lens_2} \OfType
    \Regex_1 \Or \Regex_2
    \Leftrightarrow \RegexAlt_1 \Or \RegexAlt_2
  }
\end{mathpar}
Consider finding a lens of type $\Regex_1 \Or \Regex_2 \Leftrightarrow
\RegexAlt_1 \Or \RegexAlt_2$. With only syntax-directed rules, the only lens
that can be well-typed would be an or lens.

Composition sequentially composes two lenses.
\begin{mathpar}
  \inferrule[\ComposeLensRule{}]
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \Regex_2\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3\\
  }
  {
    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Leftrightarrow \Regex_3
  }
\end{mathpar}
Composition is difficult in the context of lens synthesis. If trying to
synthesize a composition lens, one has to pull the central regular expression
``out of thin air.''

The last typing rule is type-equivalence. If two regular expressions are
star-semiring equivalent to the type of a lens, those equivalent regular
expressions also serve as the type of the lens.
\begin{mathpar}
  \inferrule[\RewriteRegexLensRule{}]
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \SSREquiv \Regex'\\
    \RegexAlt \SSREquiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }
\end{mathpar}
This rule is difficult in the context of synthesis, as it forces a search
through equivalent regular expressions.  This rule can also be applied at any
point in the derivation, which makes the search even harder.

%\begin{figure}
%  \centering
%  \begin{tikzpicture}[auto,node distance=1.5cm]
%    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};
%    
%    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
%    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
%    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };
%    
%    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
%    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};
%
%    \node[text width=2.9cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
%    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};
%    
%    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};
%    
%    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};
%    
%    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};
%    
%    
%    \path[->] (regex1.east) edge (todnfregex.north west);
%    \path[->] (regex2.east) edge (todnfregex.south west);
%    
%    \path[->] (todnfregex.north east) edge (dnfregex1.west);
%    \path[->] (todnfregex.south east) edge (dnfregex2.west);
%    
%    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
%    \path[->] (dnfregex2.east) edge (synthdnflens.south west);
%    
%    \path[->] (synthdnflens) edge (dnflens);
%    
%    \path[->] (dnflens) edge (tolens);
%    
%    \path[->] (tolens) edge (lens);
%    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
%    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
%    % Now place a relation (ID=rel1)
%    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
%    %\node (below-gen) [below=.5cm of sketch-gen] {};
%    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
%    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
%    %\node (below-compl) [below=.5cm of sketch-compl] {};
%    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
%    %% Draw an edge between rel1 and node1; rel1 and node2
%    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
%    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
%    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
%    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);
%
%    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
%    %
%    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure) (success) ,draw] (surrounding) {};
%  \end{tikzpicture}
%  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
%    \RegexAlt{}, and examples, \Examples{}, are given as input.
%    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
%    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
%    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
%    \RegexAlt{}, and \Examples{}.
%    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
%    that is equivalent to \DNFLens{}.}
%  \label{fig:schematic-diagram-synthesis}
%\end{figure}

To address the difficulties with composition and type-equivalence rules, we
synthesize lenses in an alternative language of disjunctive normal form (DNF)
lenses. DNF lenses are in pseudonormal form, containing no composition operator,
and so their synthesis never needs to pull regular expressions ``out of thin
air.'' The types of DNF lenses are pairs of regular expressions in a
pseudonormal form, DNF regular expressions. Because DNF regular expressions are
in a pseudonormal form, fewer equivalent regular expressions need to be searched
through.

In our search algorithm, we only search through equivalent regular expressions
once, before processing any syntax directed-rules. We formalize this in the
typing of DNF regular expressions by only permitting the application of
type-equivalence once, after all syntactic rules have been applied. This is
enforced by having two typing judgements: one for the ``rewriteless'' type of
the lens (meaning no type-equivalence rules were applied) and one of the
``full'' type of the lens. If $\DNFLens \OfRewritelessType \DNFRegex
\Leftrightarrow \DNFRegexAlt$, then $\DNFLens$ is a DNF lens of rewriteless type
$\DNFRegex \Leftrightarrow \DNFRegexAlt$. If $\DNFLens \OfType \DNFRegex
\Leftrightarrow \DNFRegexAlt$, then $\DNFLens$ is a DNF lens of full type
$\DNFRegex \Leftrightarrow \DNFRegexAlt$. The following rule is used to get the
full type of a DNF lens from the rewriteless type.
\begin{mathpar}
  \inferrule[\RewriteDNFRegexLensRule{}]
  {
    \DNFRegex' \StarOf{\Rewrite} \DNFRegex\\
    \DNFRegexAlt' \StarOf{\Rewrite} \DNFRegexAlt \\
    \DNFLens \OfRewritelessType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt}
  }
  {
    \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex'}{\DNFRegexAlt'}
  }
\end{mathpar}
Note that instead of using directionless equivalences,
\RewriteDNFRegexLensRule{} uses directioned rewrites. The relationship between
the two is formalized by the following theorem:
\begin{theorem}
  \label{thm:thm1}
  If $\Regex \SSREquiv \RegexAlt$, then $\ToDNFRegexOf{\Regex}
  \equiv_{\rightarrow} \ToDNFRegexOf{\RegexAlt}$, where $\ToDNFRegexOf{\Regex}$
  and $\ToDNFRegexOf{\RegexAlt}$ are $\Regex$ and $\RegexAlt$ in DNF form
  (respectively), and $\equiv_{\rightarrow}$ is the transitive, reflexive, and
  symmetric closure of $\rightarrow$.
\end{theorem}

Proving that our search procedure can generate any lens reduces to proving DNF
lenses complete with respect to our standard lens language. In particular, we
wish the prove the following theorem:
\begin{theorem}
  If $\Lens : \Regex \Leftrightarrow \RegexAlt$, then there exists a DNF lens,
  $\DNFLens$, such that $\DNFLens \OfType \ToDNFRegexOf{\Regex} \Leftrightarrow
  \ToDNFRegexOf{\RegexAlt}$ and the semantics of $\Lens$ and $\DNFLens$
  are equivalent.
\end{theorem}

We prove this propety by induction on the structure of the typing derivation.
The particular difficulty lies in the lens equivalence rule. We begin this case
below:

@@

By induction assumption, there exists $\DNFLens \OfType \ToDNFRegexOf{\Regex_1}
\Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}$, where the semantics of $\DNFLens$
are equivalent to those of $\Lens$. By inversion on the derivation of $\DNFLens
\OfType \ToDNFRegexOf{\Regex_1} \Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}$,
there exists $\DNFRegex_1$ and $\DNFRegexAlt_1$ such that:
\begin{mathpar}
  \inferrule
  {
    \ToDNFRegexOf{\Regex_1} \StarOf{\Rewrite} \DNFRegex_1\\
    \ToDNFRegexOf{\RegexAlt_1} \StarOf{\Rewrite} \DNFRegexAlt_1\\
    \DNFLens \OfRewritelessType \DNFRegex_1 \Leftrightarrow \DNFRegexAlt_1
  }
  {
    \DNFLens \OfType \ToDNFRegexOf{\Regex_1} \Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}
  }
\end{mathpar}

@@

We first show that there exists $\DNFRegex_2$ and $\DNFRegexAlt_2$ such that
$\ToDNFRegexOf{\Regex_2} \StarOf{\rightarrow} \DNFRegex_2$ and
$\ToDNFRegexOf{\DNFRegex_1} \StarOf{\rightarrow} \DNFRegex_2$ and
$\ToDNFRegexOf{\RegexAlt_2} \StarOf{\rightarrow} \DNFRegexAlt_2$ and
$\ToDNFRegexOf{\DNFRegexAlt_1} \StarOf{\rightarrow} \DNFRegexAlt_2$. To do this,
we first prove that $\StarOf{\rightarrow}$ is confluent. By
Theorem~\ref{thm:thm1}, we know that $\ToDNFRegexOf{\Regex_1} \equiv_\rightarrow
\ToDNFRegexOf{\Regex_2}$. Such $\DNFRegex_2$ and $\DNFRegexAlt_2$ are guaranteed
to exist by a corollary to confluence.

@@

After this, we have $\DNFLens \OfRewritelessType \DNFRegex_1 \Leftrightarrow
\DNFRegexAlt_1$ and $\DNFRegex_1 \StarOf{\rightarrow} \DNFRegex_2$ and
$\DNFRegexAlt_1 \StarOf{\rightarrow} \DNFRegexAlt_2$. If we can prove a
confluence-like property that would show the existance of some $\DNFLens'
\OfRewritelessType \DNFRegex_3 \Leftrightarrow \DNFRegexAlt_3$, where
$\DNFRegex_2 \StarOf{\rightarrow} \DNFRegex_3$ and $\DNFRegexAlt_2
\StarOf{\rightarrow} \DNFRegexAlt_3$, we would be done. This property is
$R$-confluence for a properly chosen $R$. This case is diagrammed in
Figure~\ref{fig:equiv-case}.

@@

\section{Related Work}

The full proof of completeness is contained in the appendix of the full version
of the original optician paper~\cite{extended-version}. The original proof of
$R$-confluence for the transitive closure of $\rightarrow$ previously required
additional assumptions. These unnecessary assumptions have been identified and
removed in this version of the proof. Future work used the proof of completeness
over our lens language to show that quotient bijective lenses are also
synthesizeable~\cite{maina+:quotient-synthesis}. Lastly, while synthesizeability
was not proven for symmetric lenses~\cite{soptician}, such a proof would likely
have proven $R$-confluence in a similar manner. Our proof used approaches
inspired by the original Church-Rosser paper~\cite{Church1936SomePO}.

This work continues a trend in making programming easier through
synthesis~\cite{gulwani2017program}. While synthesis is one approach to make
bidirectional programming easier, it is not the only approach. Recent work has
gone into building lenses without requiring the point-free combinator
style~\cite{10.1007/978-3-319-89884-1_2}. Other work has found
applicative~\cite{10.1145/2858949.2784750} and
monadic~\cite{10.1007/978-3-030-17184-1_6} approaches to compositionally
building lenses.
